---
name: Alex
surname: Groot
position: "Student Life Sciences"
# pronouns: she/her
# address: "School of Physics & Chemistry, École Normale Supérieure"
phone: +31 6 18 33 64 55
www: alexpgroot.github.io
email: "alex_groot@icloud.com"
# twitter: mariecurie
github: AlexPGroot
# linkedin: mariecurie
# date: "`r format(Sys.time(), '%B %Y')`"
output:
  vitae::awesomecv:
    page_total: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(vitae)
library(tibble)
```

# About Me

Student HBO life sciences at Hogeschool Utrecht majoring in molecular biology, transitioning into the field of data science. Having experience in R programming and data science, I am eager to use my analytical skills and programming on biological data to gain data-driven insights into processes. I also enjoy working on projects that assist in facilitating the process of analytics and research.

# Skills

- R (programming language)
- R Quarto
- R Shiny
- Data visualization (ggplot2/plotly)
- Git
- Bash
- Linux
- Microsoft Excel

# Projects

```{r}
tribble(
  ~ project, ~ subtitle, ~ skills,
  ###
  "Project Nanopore", 
  "Processing and analyzing MinION RNA-sequencing data", 
  c("FastQC", "minimap2", "Bambu", "DESeq","IGV browser"),
  ###
  "Getting started with machine learning",
  "Personal project using breast cancer dataset",
  c("Machine learning using R Tidymodels"),
  ###
  "Project genes and proteins",
  "Transfecting eukaryote cells with recombinant plasmids",
  c("SDS-PAGE", "Western-blot"),
  ###
  "Project sphingolipids",
  "Cell culture and experimentation",
  c("Working with multiple cell lines (SKNAS, HEK293T, HEPG2)", "Setting up and performing assays (MTT, Amplex Red)", "Thin Layer Chromatography (TLC)")
) %>% detailed_entries(what = subtitle, 
                       when = "", 
                       with = project, 
                       where = "", 
                       why = skills)
```
