---
name: Alex
surname: Groot
position: "Bachelor Life Sciences"
phone: +31 6 18 33 64 55
www: alexpgroot.github.io
email: "alex_groot@icloud.com"
github: AlexPGroot
linkedin: alex-groot-b10b7634b
# date: "`r format(Sys.time(), '%B %Y')`"
profilepic: "$HOME/Pictures/IMG_1306.JPEG"
output:
  vitae::awesomecv:
    page_total: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(vitae)
library(tibble)
```

# About Me

Life Sciences graduate with a strong interest in bioinformatics, comparative genomics, and data science. During my graduation internship at Amsterdam UMC, Leptospirosis Reference Centre, I worked on pangenome analyses of 1440 _Leptospira_ genomes, investigating genetic variation and host adaptation using computational approaches. Skilled in R programming, data analysis, and scientific workflows, I enjoy applying quantitative methods to \newline biological questions and developing tools that support efficient and reproducible research.

# Work experience and internships

```{r}
tribble(
  ~ Skills, ~ Year, ~ Where, ~ Experience,
  "Bioinformatics, Pangenome, Phylogeny, HPC, Virulence factors, Vaccine target screening", "2025-2026", "Amsterdam UMC, Leptospirosis Reference Centre", "Graduation internship",
  "Plasmid/primer design, Biotechnology, CrisprCas", "2020-2021", "Leiden UMC", "Internship"
) %>%
  detailed_entries(Skills, Year, Where, Experience)
```

# Education

```{r}
tribble(
  ~ Degree, ~ Year, ~ Institution, ~ Where,
  "Bachelor of Science, Biologie en Medisch Laboratoriumonderzoek (Life Sciences)", "2018-2026", "Hogeschool Utrecht", "Utrecht, The Netherlands"
) %>%
  detailed_entries(Degree, Year, Institution, Where)
```

# Skills

## Bioinformatics & Genomics

* Comparative genomics and pangenomics
* Phylogenetic analysis
* Genome annotation and functional analysis
* Long-read sequencing data analysis (Oxford Nanopore)
* R (tidyverse, tidymodels)
* Data visualization (ggplot2, plotly)
* Machine learning (R Tidymodel)

## General Tools

* Linux server
* Bash
* Git
* High Performance Computing (HPC)
* LaTeX

## Laboratory Techniques

* Cell culture
* Recombinant DNA techniques
* SDS-PAGE
* Western blotting
* Thin Layer Chromatography (TLC)
* MTT and Amplex Red assays

\newpage
# Projects

```{r}
tribble(
  ~ project, ~ subtitle, ~ skills,
  ###
  "Pangenome investigation of Leptospira genus and insight into the epidemiology of Leptospirosis",
  "One year graduation internship at Amsterdam UMC Leptospirosis reference centre",
  c(
  "Pangenome analysis (PPanGGOLiN)",
  "Genome-wide association studies (Scoary)",
  "Phylogenetic analysis (IQ-TREE)",
  "Functional annotation (COG categories)",
  "High Performance Computing (HPC)",
  "Data visualization and statistical analysis"
),
  ###
  "Project Nanopore", 
  "Processing and analyzing MinION RNA-sequencing data", 
  c("FastQC", "minimap2", "Bambu", "DESeq","IGV browser"),
  ###
  "Project genes and proteins",
  "Transfecting eukaryote cells with recombinant plasmids",
  c("SDS-PAGE", "Western-blot"),
  ###
  "Project sphingolipids",
  "Cell culture and experimentation",
  c("Working with multiple cell lines (SKNAS, HEK293T, HEPG2)", "Setting up and performing assays (MTT, Amplex Red)", "Thin Layer Chromatography (TLC)")
) %>% detailed_entries(what = subtitle, 
                       when = "", 
                       with = project, 
                       where = "", 
                       why = skills)
```
